### 2023-04-06 result_sandbox.Rmd
Switching from feat_id to feature and height to area during cleanup

---
title: "MS_metrics results"
author: "William Kumler"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
FT2040_features <- read_csv("made_data_FT2040/features_extracted.csv")
MS3000_features <- read_csv("made_data_MS3000/features_extracted.csv")
```

Can we simply train up a model on one dataset and apply it to another, i.e. how much does our performance suffer when switching to a different dataset?

Test with MESOSCOPE and Falkor, two fully-labeled feature-based datasets. Train up with Falkor and test on MESOSCOPE, train up on MESOSCOPE and test on Falkor, compare model outputs (confusion matrices and model estimates).

```{r Confusion matrices and figure}
dataset_versions <- c("FT2040_features", "MS3000_features")
traintest_gp <- lapply(dataset_versions, function(train_set){
  full_model <- get(train_set) %>% 
    select(-feature) %>%
    filter(feat_class%in%c("Good", "Bad")) %>%
    mutate(feat_class=feat_class=="Good") %>%
    glm(formula=feat_class~., family = binomial)
  lapply(dataset_versions, function(test_set){
    get(test_set) %>%
      select(-feature) %>%
      filter(feat_class%in%c("Good", "Bad")) %>%
      mutate(pred_prob=predict(object=full_model,newdata = ., 
                               type = "response")) %>%
      mutate(pred_class_50=ifelse(pred_prob>0.5, "Good", "Bad")) %>%
      mutate(pred_class_99=ifelse(pred_prob>0.99, "Good", "Bad")) %>%
      mutate(pred_class_01=ifelse(pred_prob>0.01, "Good", "Bad")) %>%
      select(feat_class, starts_with("pred_class")) %>%
      mutate(train_test=paste(train_set, test_set, sep = "-")) %>%
      mutate(train_test=str_remove_all(train_test, "_features")) %>%
      mutate(cross_type=ifelse(train_set==test_set, "Within", "Cross"))
  }) %>% bind_rows()
}) %>%
  bind_rows() %>%
  pivot_longer(cols = starts_with("pred_class"), 
               names_to = "threshold", 
               values_to = "pred_class") %>%
  mutate(error_type=case_when(
    feat_class=="Bad" & pred_class=="Bad" ~ "TN",
    feat_class=="Bad" & pred_class=="Good" ~ "FP",
    feat_class=="Good" & pred_class=="Bad" ~ "FN",
    feat_class=="Good" & pred_class=="Good" ~ "TP"
  )) %>%
  group_by(cross_type, train_test, threshold) %>%
  count(error_type) %>%
  mutate(perc=round(n/sum(n)*100)) %>%
  ungroup() %>%
  complete(error_type, train_test, threshold, fill = list(n=0, perc=0)) %>%
  mutate(cross_type=ifelse(is.na(cross_type), "Within", cross_type)) %>%
  # Rename things to look nice in ggplot
  mutate(error_type=paste0("%", error_type)) %>%
  mutate(error_type=factor(error_type, paste0("%", c("TN", "TP", "FN", "FP")))) %>%
  mutate(train_test=str_replace_all(train_test, "FT2040", "Falkor")) %>%
  mutate(train_test=str_replace_all(train_test, "MS3000", "MESO")) %>%
  mutate(threshold=as.numeric(str_extract(threshold, "\\d+"))) %>%
  mutate(threshold=paste("Threshold:", threshold/100)) %>%
  mutate(cross_type=paste(cross_type, "dataset", sep = "-")) %>% 
  mutate(cross_type=factor(cross_type, levels=c("Within-dataset", "Cross-dataset"))) %>%
  arrange(train_test) %>% 
  mutate(train_test=factor(train_test, levels=unique(train_test))) %>%
  # Render the ggplot
  ggplot(aes(x=error_type, y=perc)) +
  geom_col(aes(fill=train_test), position = position_dodge()) +
  geom_text(aes(label=n, color=train_test), 
            position = position_dodge(width = 1), vjust=-0.2) +
  facet_grid(threshold~cross_type, scales = "free_x") +
  scale_y_continuous(limits = c(0, 100)) +
  scale_color_discrete("Train-test", aesthetics=c("color", "fill")) +
  theme_bw() +
  theme(axis.title = element_blank())
xcms_gp <- list(Falkor=FT2040_features, MESO=MS3000_features) %>%
  bind_rows(.id = "cruise") %>%
  filter(feat_class%in%c("Good", "Bad")) %>%
  group_by(cruise) %>%
  count(feat_class) %>%
  mutate(perc=round(n/sum(n)*100)) %>%
  ungroup() %>%
  add_row(cruise=c("Falkor", "MESO"), feat_class="Bad", perc=NA) %>%
  add_row(cruise=c("Falkor", "MESO"), feat_class="Good", perc=NA) %>%
  arrange(cruise, feat_class, n) %>%
  mutate(pred_class=rep(c("Good", "Bad"), length.out=8)) %>%
  mutate(error_type=case_when(
    feat_class=="Bad" & pred_class=="Bad" ~ "TN",
    feat_class=="Bad" & pred_class=="Good" ~ "FP",
    feat_class=="Good" & pred_class=="Bad" ~ "FN",
    feat_class=="Good" & pred_class=="Good" ~ "TP"
  )) %>%
  mutate(error_type=paste0("%", error_type)) %>%
  mutate(error_type=factor(error_type, paste0("%", c("TN", "TP", "FN", "FP")))) %>%
  ggplot(aes(x=error_type, y=perc)) +
  geom_col(aes(fill=cruise), position = "dodge") +
  geom_text(aes(label=n, color=cruise), 
            position = position_dodge(width = 1), vjust=-0.2) +
  scale_y_continuous(limits = c(0, 100)) +
  scale_fill_discrete("Dataset", aesthetics=c("color", "fill")) +
  theme_bw() +
  theme(legend.background = element_rect(color="black"),
        axis.title = element_blank()) +
  # theme(legend.position = c(0.5, 1),
  #       legend.justification = c(0.5, 1), 
  #       legend.direction = "horizontal") +
  theme(legend.position = c(0, 1),
        legend.justification = c(0, 1),
        legend.direction = "vertical")
plot(gridExtra::arrangeGrob(
  xcms_gp + ggtitle("A. XCMS output"), 
  traintest_gp + ggtitle("B. Performance of logistic model"), 
  layout_matrix = matrix(c(1,2,2,2), nrow = 1)))
```

```{r Confusion matrices and figure, v2 clean}
dataset_versions <- c("FT2040_features", "MS3000_features")
confusion_data <- lapply(dataset_versions, function(train_set){
  full_model <- get(train_set) %>% 
    select(-feature) %>%
    filter(feat_class%in%c("Good", "Bad")) %>%
    mutate(feat_class=feat_class=="Good") %>%
    glm(formula=feat_class~., family = binomial)
  lapply(dataset_versions, function(test_set){
    get(test_set) %>%
      select(-feature) %>%
      filter(feat_class%in%c("Good", "Bad")) %>%
      mutate(pred_prob=predict(object=full_model,newdata = ., 
                               type = "response")) %>%
      mutate(pred_class=ifelse(pred_prob>0.5, "Good", "Bad")) %>%
      select(feat_class, starts_with("pred_class")) %>%
      mutate(train_test=paste(train_set, test_set, sep = "-")) %>%
      mutate(train_test=str_remove_all(train_test, "_features")) %>%
      mutate(cross_type=ifelse(train_set==test_set, "Within", "Cross"))
  }) %>% bind_rows()
}) %>%
  bind_rows() %>%
  mutate(error_type=case_when(
    feat_class=="Bad" & pred_class=="Bad" ~ "TN",
    feat_class=="Bad" & pred_class=="Good" ~ "FP",
    feat_class=="Good" & pred_class=="Bad" ~ "FN",
    feat_class=="Good" & pred_class=="Good" ~ "TP"
  )) %>%
  group_by(cross_type, train_test) %>%
  count(error_type) %>%
  mutate(perc=round(n/sum(n)*100)) %>%
  ungroup() %>%
  # Rename things to look nice in ggplot
  mutate(error_type=paste0("%", error_type)) %>%
  mutate(error_type=factor(error_type, paste0("%", c("TN", "TP", "FN", "FP")))) %>%
  mutate(train_test=str_replace_all(train_test, "FT2040", "Falkor")) %>%
  mutate(train_test=str_replace_all(train_test, "MS3000", "MESO")) %>%
  mutate(cross_type=paste(cross_type, "dataset", sep = "-")) %>% 
  mutate(cross_type=factor(cross_type, levels=c("Within-dataset", "Cross-dataset"))) %>%
  arrange(train_test) %>% 
  mutate(train_test=factor(train_test, levels=unique(train_test)))

within_gp <- confusion_data %>%
  filter(cross_type=="Within-dataset") %>%
  ggplot(aes(x=error_type, y=perc)) +
  geom_col(aes(fill=train_test), position = position_dodge()) +
  geom_text(aes(label=n, color=train_test), 
            position = position_dodge(width = 1), vjust=-0.2) +
  scale_y_continuous(limits = c(0, 100)) +
  scale_color_discrete("Train-test", aesthetics=c("color", "fill")) +
  theme_bw() +
  theme(axis.title = element_blank(),
      legend.background = element_rect(color="black"),
      legend.position = c(1, 1),
      legend.justification = c(1, 1),
      legend.direction = "vertical",
      axis.text.y = element_blank())
cross_gp <- confusion_data %>%
  filter(cross_type=="Cross-dataset") %>%
  ggplot(aes(x=error_type, y=perc)) +
  geom_col(aes(fill=train_test), position = position_dodge()) +
  geom_text(aes(label=n, color=train_test), 
            position = position_dodge(width = 1), vjust=-0.2) +
  scale_y_continuous(limits = c(0, 100)) +
  scale_color_discrete("Train-test", aesthetics=c("color", "fill")) +
  theme_bw() +
  theme(axis.title = element_blank(),
      legend.background = element_rect(color="black"),
      legend.position = c(1, 1),
      legend.justification = c(1, 1),
      legend.direction = "vertical",
      axis.text.y = element_blank())
xcms_gp <- list(Falkor=FT2040_features, MESO=MS3000_features) %>%
  bind_rows(.id = "cruise") %>%
  filter(feat_class%in%c("Good", "Bad")) %>%
  group_by(cruise) %>%
  count(feat_class) %>%
  mutate(perc=round(n/sum(n)*100)) %>%
  ungroup() %>%
  add_row(cruise=c("Falkor", "MESO"), feat_class="Bad", perc=NA) %>%
  add_row(cruise=c("Falkor", "MESO"), feat_class="Good", perc=NA) %>%
  arrange(cruise, feat_class, n) %>%
  mutate(pred_class=rep(c("Good", "Bad"), length.out=8)) %>%
  mutate(error_type=case_when(
    feat_class=="Bad" & pred_class=="Bad" ~ "TN",
    feat_class=="Bad" & pred_class=="Good" ~ "FP",
    feat_class=="Good" & pred_class=="Bad" ~ "FN",
    feat_class=="Good" & pred_class=="Good" ~ "TP"
  )) %>%
  mutate(error_type=paste0("%", error_type)) %>%
  mutate(error_type=factor(error_type, paste0("%", c("TN", "TP", "FN", "FP")))) %>%
  ggplot(aes(x=error_type, y=perc)) +
  geom_col(aes(fill=cruise), position = "dodge") +
  geom_text(aes(label=n, color=cruise), 
            position = position_dodge(width = 1), vjust=-0.2) +
  scale_y_continuous(limits = c(0, 100)) +
  scale_fill_discrete("Dataset", aesthetics=c("color", "fill")) +
  theme_bw() +
  theme(axis.title = element_blank(),
        legend.background = element_rect(color="black"),
        legend.position = c(0, 1),
        legend.justification = c(0, 1),
        legend.direction = "vertical")
plot(gridExtra::arrangeGrob(
  xcms_gp + ggtitle("A. Default XCMS output\n    (No model fitting)"), 
  within_gp + ggtitle("B. Performance within\n     a given dataset"), 
  cross_gp + ggtitle("C. Performance across\n     datasets"), 
  layout_matrix = matrix(c(1,2,3), nrow = 1)))
```

```{r Thresholding demo}
meso_full_model <- MS3000_features %>% 
  select(-feature) %>%
  filter(feat_class%in%c("Good", "Bad")) %>%
  mutate(feat_class=feat_class=="Good") %>%
  glm(formula=feat_class~., family = binomial)

plot_breaks <- c(0.01, 0.1, 0.5, 0.9, 0.99)
base_plot <- MS3000_features %>%
  mutate(pred_prob=predict(object=meso_full_model,newdata = ., type = "response")) %>%
  mutate(cut_prob=cut(pred_prob, breaks = c(0, plot_breaks, 1),include.lowest = TRUE)) %>%
  mutate(feat_class=factor(feat_class, levels=c("Bad", "Stans only", "Meh", "Good"))) %>%
  ggplot() +
  geom_bar(aes(x=cut_prob, fill=feat_class), 
           position =position_dodge(preserve = "single")) +
  facet_wrap(~cut_prob, nrow=1, scales = "free_x") +
  scale_y_log10() +
  scale_fill_discrete("Manual feature classification") +
  labs(x="Predicted probability (binned) of being a good peak", y="Count") +
  theme_bw() +
  theme(legend.position = "top", legend.justification = "left") +
  guides(fill=guide_legend(title.position="top"))

threshold_plot <- lapply(plot_breaks, function(threshold){
  MS3000_features %>%
    filter(feat_class%in%c("Good", "Bad")) %>%
    mutate(pred_prob=predict(object=meso_full_model,newdata = ., type = "response")) %>%
    mutate(pred_class=ifelse(pred_prob>threshold, "Good", "Bad")) %>%
    mutate(error_type=case_when(
      feat_class=="Bad" & pred_class=="Bad" ~ "TN",
      feat_class=="Bad" & pred_class=="Good" ~ "FP",
      feat_class=="Good" & pred_class=="Bad" ~ "FN",
      feat_class=="Good" & pred_class=="Good" ~ "TP"
    )) %>%
    count(error_type) %>%
    mutate(perc=round(n/sum(n)*100)) %>%
    ungroup() %>%
    mutate(threshold=threshold)
}) %>%
  bind_rows() %>%
  mutate(error_type=paste0("%", error_type)) %>%
  mutate(error_type=factor(error_type, paste0("%", c("TN", "TP", "FN", "FP")))) %>%
  mutate(threshold=paste("Threshold:", threshold)) %>%
  ggplot(aes(x=error_type, y=perc)) +
  geom_col() +
  geom_text(aes(label=n), vjust=-0.5) +
  facet_wrap(~threshold, nrow = 1) +
  theme_bw() +
  theme(axis.title = element_blank()) +
  scale_y_continuous(limits = c(0, 100))

layout_mat <- matrix(c(c(1,1,1,1,1,1,1,1,1,1),c(NA,2,2,2,2,2,2,2,2,NA)), nrow = 2, byrow = TRUE)
plot(gridExtra::arrangeGrob(base_plot, threshold_plot, layout_matrix = layout_mat))
```

```{r Backup confusion matrix visuals, eval=FALSE}
full_cmat <- rbind(
  rbind(
    data.frame(f2f_cmat) %>% mutate(tested_on="falkor"),
    data.frame(f2m_cmat) %>% mutate(tested_on="mesoscope")
  ) %>% mutate(trained_on="falkor"),
  rbind(
    data.frame(m2m_cmat) %>% mutate(tested_on="mesoscope"),
    data.frame(m2f_cmat) %>% mutate(tested_on="falkor")
  ) %>% mutate(trained_on="mesoscope")
)

library(ggplot2)
full_cmat %>%
  mutate(predicted=paste("Predicted to\nbe", predicted)) %>%
  mutate(predicted=factor(predicted, levels=c("Predicted to\nbe Good", "Predicted to\nbe Bad"))) %>%
  mutate(actual=paste("Actually", actual)) %>%
  mutate(tested_on=paste("Tested on", tested_on)) %>%
  mutate(trained_on=paste("Trained on", trained_on)) %>%
  ggplot() +
  geom_text(aes(x=actual, y=predicted, label=Freq)) +
  facet_grid(tested_on~trained_on, switch = "y") +
  scale_x_discrete(position = "top") + 
  theme_bw() +
  theme(strip.placement = "outside", axis.title = element_blank(),
        axis.text.y = element_text(angle=90, hjust = 0.5, vjust = 0),
        strip.background = element_rect(fill = "white"))

library(flextable)
full_cmat %>%
  arrange(trained_on, tested_on, predicted, actual) %>%
  select(`Trained on`=trained_on, `Tested on`=tested_on, Predicted=predicted, 
         Actual=actual, Count=Freq) %>%
  flextable() %>%
  merge_v(j = "Trained on") %>%
  merge_v(j = "Tested on") %>%
  merge_v(j = "Predicted") %>%
  align(align = "center") %>%
  rotate(i = 1:16, j=1:2, rotation = "btlr") %>%
  theme_box()

library(gt)
cmat_to_df <- function(cmat){
  outmat <- matrix(c(
    "", "Bad", "Good",
    "Bad", cmat[1,1], cmat[1,2],
    "Good", cmat[2,1], cmat[2,2]
  ), nrow = 3, ncol = 3)
  colnames(outmat) <- c("Predicted", "Actual", "Actual2")
  outmat
}
as.data.frame(rbind(
  rbind(cmat_to_df(f2f_cmat), cmat_to_df(f2m_cmat)),
  rbind(cmat_to_df(m2m_cmat), cmat_to_df(m2f_cmat)
  ))) %>%
  gt() %>%
  tab_row_group(
    label="Trained on Mesoscope, tested on Falkor", rows = 10:12
  ) %>%
  tab_row_group(
    label="Trained on Mesoscope, tested on Mesoscope", rows = 7:9
  ) %>%
  tab_row_group(
    label="Trained on Falkor, tested on Mesoscope", rows = 4:6
  ) %>%
  tab_row_group(
    label="Trained on Falkor, tested on Falkor", rows = 1:3
  ) %>%
  tab_options(column_labels.hidden = TRUE)
```

Model estimates:

```{r Model comparison}
falkor_full_model <- FT2040_features %>% 
  select(-feature) %>%
  filter(feat_class%in%c("Good", "Bad")) %>%
  mutate(feat_class=feat_class=="Good") %>%
  glm(formula=feat_class~., family = binomial)
meso_full_model <- MS3000_features %>% 
  select(-feature) %>%
  filter(feat_class%in%c("Good", "Bad")) %>%
  mutate(feat_class=feat_class=="Good") %>%
  glm(formula=feat_class~., family = binomial)

good_params <-list(falkor=broom::tidy(falkor_full_model),
                   meso=broom::tidy(meso_full_model)) %>%
  bind_rows(.id="dataset") %>%
  group_by(term) %>%
  mutate(keep_term=all(p.value<0.05)) %>%
  filter(keep_term)

good_params %>%
  ggplot(aes(x=term, color=dataset)) +
  geom_hline(yintercept = 0, color="black") +
  geom_point(aes(y=estimate), position = position_dodge(width = 0.5), size=3) +
  geom_errorbar(aes(ymin=estimate-2*std.error, ymax=estimate+2*std.error), 
                position = position_dodge(width = 0.5), linewidth=1) +
  facet_wrap(~term, scales = "free") +
  labs(x=NULL, y="Model parameter estimate") +
  theme_bw() +
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank())

bind_rows(falkor=FT2040_features, meso=MS3000_features, .id = "dataset") %>%
  select(dataset, all_of(unique(good_params$term)), feat_class) %>%
  filter(feat_class%in%c("Good", "Bad")) %>%
  mutate(feat_class=ifelse(feat_class=="Good", 1, 0)) %>%
  mutate(feat_class_jitter=feat_class+rnorm(nrow(.), sd = 0.02)) %>%
  pivot_longer(-c(dataset, feat_class, feat_class_jitter), names_to = "param") %>%
  slice_sample(prop = 1) %>%
  ggplot(aes(x=value, y=feat_class)) +
  geom_point(aes(y=feat_class_jitter, color=dataset), alpha=0.1) +
  geom_smooth(aes(group=dataset), formula="y~x", method = "glm",
              method.args=list(family=binomial),
              se=FALSE, linewidth=2, color="white") +
  geom_smooth(aes(color=dataset), formula="y~x", method = "glm",
              method.args=list(family=binomial)) +
  facet_wrap(~param, scales = "free_x")

bind_rows(list(Falkor=FT2040_features, MESOSCOPE=MS3000_features), .id = "cruise") %>%
  ggplot() + 
  geom_histogram(aes(x=med_cor, fill=feat_class), bins=40) +
  facet_wrap(~cruise, ncol = 1, scales = "free_y")
```

Repeat the above with a reduced set of predictors (just the "best" ones). Maybe also with just XCMS parameters?

```{r feature selection and selection testing}
raw_data_params <- c("med_cor", "med_SNR")
xcms_params <- c("mean_mz", "sd_ppm", "mean_rt", "sd_rt", "mean_pw", 
                 "sd_pw", "log_mean_area", "log_sd_area", "sn", 
                 "f", "scale", "lmin", "feat_npeaks", "n_found", 
                 "samps_found", "stans_found")
all_params <- c("mean_mz", "sd_ppm", "mean_rt", "sd_rt", "mean_pw", "sd_pw",
                "log_mean_area", "log_sd_area", "sn", "f", "scale", 
                "lmin", "feat_npeaks", "n_found", "samps_found", "stans_found", 
                "med_cor", "med_SNR", "med_missed_scans", "smp_to_blk", 
                "smp_to_std", "shape_cor", "area_cor", "feat_class")
lst(all_params, raw_data_params, xcms_params) %>%
  imap(function(param_selection, param_name){
    lapply(dataset_versions, function(train_set){
      full_model <- get(train_set) %>% 
        select(feat_class, all_of(param_selection)) %>%
        filter(feat_class%in%c("Good", "Bad")) %>%
        mutate(feat_class=feat_class=="Good") %>%
        glm(formula=feat_class~., family = binomial)
      lapply(dataset_versions, function(test_set){
        get(test_set) %>%
          filter(feat_class%in%c("Good", "Bad")) %>%
          mutate(pred_prob=predict(object=full_model,newdata = ., 
                                   type = "response")) %>%
          mutate(pred_class=ifelse(pred_prob>0.5, "Good", "Bad")) %>%
          select(feat_class, starts_with("pred_class")) %>%
          mutate(train_test=paste(train_set, test_set, sep = "-")) %>%
          mutate(train_test=str_remove_all(train_test, "_features")) %>%
          mutate(cross_type=ifelse(train_set==test_set, "Within", "Cross"))
      }) %>% bind_rows()
    }) %>% bind_rows() %>% mutate(which_params=param_name)
  }) %>% bind_rows() %>%
  mutate(error_type=case_when(
    feat_class=="Bad" & pred_class=="Bad" ~ "TN",
    feat_class=="Bad" & pred_class=="Good" ~ "FP",
    feat_class=="Good" & pred_class=="Bad" ~ "FN",
    feat_class=="Good" & pred_class=="Good" ~ "TP"
  )) %>%
  group_by(cross_type, train_test, which_params) %>%
  count(error_type) %>%
  mutate(perc=round(n/sum(n)*100)) %>%
  ungroup() %>%
  # Rename things to look nice in ggplot
  mutate(error_type=paste0("%", error_type)) %>%
  mutate(error_type=factor(error_type, paste0("%", c("TN", "TP", "FN", "FP")))) %>%
  mutate(train_test=str_replace_all(train_test, "FT2040", "Falkor")) %>%
  mutate(train_test=str_replace_all(train_test, "MS3000", "MESO")) %>%
  mutate(cross_type=paste(cross_type, "dataset", sep = "-")) %>% 
  mutate(cross_type=factor(cross_type, levels=c("Within-dataset", "Cross-dataset"))) %>%
  arrange(train_test) %>% 
  mutate(train_test=factor(train_test, levels=unique(train_test))) %>%
  # Render the ggplot
  ggplot(aes(x=error_type, y=perc)) +
  geom_col(aes(fill=train_test), position = position_dodge()) +
  geom_text(aes(label=n, color=train_test), 
            position = position_dodge(width = 1), vjust=-0.2) +
  facet_grid(which_params~cross_type, scales = "free_x") +
  scale_y_continuous(limits = c(0, 100)) +
  scale_color_discrete("Train-test", aesthetics=c("color", "fill")) +
  theme_bw() +
  theme(axis.title = element_blank())
```

Should we include the subsetting analysis I did where we run the models on a subset of the data to see how large of a sample size we need to reach a stable model?

```{r model stability sample size subsetting}
# Full model
meso_full_model <- MS3000_features %>% 
  select(-feature) %>%
  filter(feat_class%in%c("Good", "Bad")) %>%
  mutate(feat_class=feat_class=="Good") %>%
  glm(formula=feat_class~., family = binomial) %>%
  broom::tidy()
set.seed(123)
meso_reps <- replicate(20, {
  meso_full_model <- MS3000_features %>% 
    slice_sample(prop = 0.5) %>%
    select(-feature) %>%
    filter(feat_class%in%c("Good", "Bad")) %>%
    mutate(feat_class=feat_class=="Good") %>%
    glm(formula=feat_class~., family = binomial) %>%
    broom::tidy()
}, simplify = FALSE) %>%
  bind_rows()
ggplot() +
  geom_vline(xintercept = 0) +
  geom_point(aes(x=estimate, y=term), data = meso_full_model, size=3) +
  geom_errorbar(aes(xmin=estimate-std.error*2, xmax=estimate+std.error*2, y=term), 
                data = meso_full_model, linewidth=1) +
  geom_point(aes(x=estimate, y=term), data = meso_reps, color="red") +
  facet_wrap(~term, scales = "free") +
  theme(axis.text.y=element_blank())

# Reduced model
meso_min_model <- MS3000_features %>% 
  select(feat_class, med_cor, med_SNR) %>%
  filter(feat_class%in%c("Good", "Bad")) %>%
  mutate(feat_class=feat_class=="Good") %>%
  glm(formula=feat_class~., family = binomial) %>%
  broom::tidy()
set.seed(123)
meso_reps <- replicate(20, {
  meso_min_model <- MS3000_features %>% 
    slice_sample(prop = 0.5) %>%
    select(feat_class, med_cor, med_SNR) %>%
    filter(feat_class%in%c("Good", "Bad")) %>%
    mutate(feat_class=feat_class=="Good") %>%
    glm(formula=feat_class~., family = binomial) %>%
    broom::tidy()
}, simplify = FALSE) %>%
  bind_rows()
ggplot() +
  geom_vline(xintercept = 0) +
  geom_point(aes(x=estimate, y=term), data = meso_min_model, size=3) +
  geom_errorbar(aes(xmin=estimate-std.error*2, xmax=estimate+std.error*2, y=term), 
                data = meso_min_model, linewidth=1) +
  geom_point(aes(x=estimate, y=term), data = meso_reps, color="red") +
  facet_wrap(~term, scales = "free") +
  theme(axis.text.y=element_blank())
```

Clearly, we can't just hand over a model if we want a level of accuracy beyond this. However, labeling things manually one-at-a-time sucks. Can we group/aggregate peaks to label them in clusters?

Yes, using PCA pixelpicking things.

```{r visualize PCA pixelpicking}
library(RaMS)
library(xcms)
dataset_version <- "FT2040"
output_folder <- paste0("made_data_", dataset_version, "/")
msnexp_filled <- readRDS(paste0(output_folder, "msnexp_filled.rds"))
feature_centers <- featureDefinitions(msnexp_filled) %>%
  as.data.frame() %>%
  select(mzmed, rtmed) %>%
  rownames_to_column("feature") %>%
  mutate(rtmed=rtmed/60) %>%
  filter(mzmed%between%c(117, 120))
msdata <- readRDS(paste0(output_folder, "msdata.rds"))
pixeldemo_msdata <- msdata$EIC2[mz%between%c(117, 120)]
interp_dt <- mapply(function(mzmed_i, rtmed_i, feature_i){
  interp_range <- rtmed_i+c(-0.5, 0.5)
  interp_points <- seq(interp_range[1], interp_range[2], length.out=50)
  pixeldemo_msdata[mz%between%pmppm(mzmed_i)] %>%
    split(.$filename) %>%
    lapply(function(eic_file){
      if(nrow(eic_file)>2){
        setNames(approx(eic_file$rt, eic_file$int, xout=interp_points), c("rt", "int"))
      } else {
        data.frame(rt=numeric(), int=numeric())
      }
    }) %>%
    bind_rows(.id="filename") %>%
    mutate(feature=feature_i)
  }, feature_centers$mzmed, feature_centers$rtmed, feature_centers$feature, 
  SIMPLIFY = FALSE) %>%
  bind_rows()
interp_dt %>%
  ggplot() +
  geom_line(aes(x=rt, y=int, group=filename)) +
  facet_wrap(~feature, scales="free") +
  theme(axis.text.y = element_blank(), axis.ticks.y=element_blank())

interp_scaled <- interp_dt %>%
  group_by(feature, filename) %>%
  mutate(int=int/max(int)) %>%
  mutate(rt=rank(rt)) %>%
  ungroup() 
  # complete(feature, filename, rt, fill = list(int=0))
interp_scaled %>%
  ggplot() +
  geom_line(aes(x=rt, y=int, color=int, group=filename)) +
  facet_wrap(~feature, scales="free") +
  scale_color_viridis_c()
interp_scaled %>%
  ggplot() +
  geom_tile(aes(x=rt, y=filename, fill=int)) +
  facet_wrap(~feature) +
  scale_fill_viridis_c() +
  theme(axis.text = element_blank(), axis.ticks = element_blank())
pcaoutput <- interp_scaled %>%
  ungroup() %>%
  pivot_wider(names_from=feature, values_from = int) %>%
  select(which(colSums(is.na(.))==0)) %>%
  arrange(filename, rt) %>%
  select(-rt, -filename) %>%
  data.matrix() %>%
  prcomp()

library(ggrepel)
pcaoutput$rotation %>%
  as.data.frame() %>%
  rownames_to_column("feature") %>%
  ggplot() +
  geom_text_repel(aes(x=PC1, y=PC2, label=feature), max.overlaps = Inf, 
                  min.segment.length = Inf) +
  theme_bw()

pcapercs <- pcaoutput$sdev^2*100
ggplot() + 
  geom_col(aes(x=seq_along(pcapercs), y=pcapercs)) +
  labs(y="% variance explained", x="PC #") +
  theme_bw() +
  scale_y_continuous(expand = c(0.01, 0.02)) +
  scale_x_continuous(breaks = seq_along(pcapercs), expand = c(0.02, 0))

cbind(distinct(interp_scaled, filename, rt), -pcaoutput$x[,1:4]) %>% 
  pivot_longer(starts_with("PC"), names_to = "PC", values_to = "int") %>%
  ggplot() +
  geom_tile(aes(x=rt, y=filename, fill=int)) +
  facet_wrap(~PC) +
  scale_fill_viridis_c() +
  theme(axis.text = element_blank(), axis.ticks = element_blank(),
        axis.title = element_blank())
```

```{r full PCA pixelpicking}
dataset_version <- "FT2040"
output_folder <- paste0("made_data_", dataset_version, "/")
# msnexp_filled <- readRDS(paste0(output_folder, "msnexp_filled.rds"))
# feature_centers <- featureDefinitions(msnexp_filled) %>%
#   as.data.frame() %>%
#   select(mzmed, rtmed) %>%
#   rownames_to_column("feature") %>%
#   mutate(rtmed=rtmed/60)
# msdata <- readRDS(paste0(output_folder, "msdata.rds"))
# interp_dt <- mapply(function(mzmed_i, rtmed_i, feature_i){
#   interp_range <- rtmed_i+c(-0.5, 0.5)
#   interp_points <- seq(interp_range[1], interp_range[2], length.out=50)
#   msdata$EIC2[mz%between%pmppm(mzmed_i)] %>%
#     split(.$filename) %>%
#     lapply(function(eic_file){
#       if(nrow(eic_file)>2){
#         setNames(approx(eic_file$rt, eic_file$int, xout=interp_points), c("rt", "int"))
#       } else {
#         data.frame(rt=numeric(), int=numeric())
#       }
#     }) %>%
#     bind_rows(.id="filename") %>%
#     mutate(feature=feature_i)
#   }, feature_centers$mzmed, feature_centers$rtmed, feature_centers$feature, 
#   SIMPLIFY = FALSE) %>%
#   bind_rows()
# saveRDS(interp_dt, file = paste0(output_folder, "interp_dt.rds"))
interp_dt <- readRDS(file = paste0(output_folder, "interp_dt.rds"))
interp_scaled <- interp_dt %>%
  group_by(feature, filename) %>%
  mutate(int=int/max(int)) %>%
  mutate(rt=rank(rt)) %>%
  ungroup()
pcaoutput <- interp_scaled %>%
  ungroup() %>%
  pivot_wider(names_from=feature, values_from = int) %>%
  select(which(colSums(is.na(.))==0)) %>%
  arrange(filename, rt) %>%
  select(-rt, -filename) %>%
  data.matrix() %>%
  prcomp()
pcaoutput$rotation %>%
  as.data.frame() %>%
  rownames_to_column("feature") %>%
  ggplot() +
  geom_text(aes(x=PC1, y=PC2, label=feature)) +
  theme_bw() +
  coord_fixed()

# Currently broken?
pc_features <- FT2040_features %>%
  select(feature, feat_class) %>%
  filter(feature%in%rownames(pcaoutput$rotation)) %>%
  cbind(pcaoutput$rotation[,c("PC1", "PC2")])

ggplot(pc_features) +
  geom_text(aes(x=PC1, y=PC2, color=feat_class, label=feature)) +
  theme_bw() +
  theme(legend.position = "none") +
  coord_fixed()

pc_circled <- pc_features %>%
  mutate(quick_class=ifelse(sqrt(PC1^2+PC2^2)<0.02, "Bad", "Unclassified")) %>%
  mutate(quick_class=ifelse(sqrt((PC1-0.08)^2+(PC2-0.05)^2)<0.03, 
                           "Good", quick_class))
pc_circled %>%
  mutate(plot_color=ifelse(quick_class=="Unclassified", NA, feat_class)) %>%
  ggplot() +
  geom_point(aes(x=PC1, y=PC2, color=plot_color)) +
  ggforce::geom_circle(aes(x0=0, y0=0, r=0.02), n = 36, linewidth=1,
                       color=scales::hue_pal()(4)[1]) +
  ggforce::geom_circle(aes(x0=0.08, y0=0.05, r=0.03), n = 36, linewidth=1,
                       color=scales::hue_pal()(4)[2]) +
  theme_bw() +
  theme(legend.position = "none") +
  coord_fixed()
```

Prove this with confusion matrices showing very few false pos/neg:

```{r PCA pixelpicking confusion matrix}
table(pc_circled$feat_class, pc_circled$quick_class)
```

How does a model trained on just the best/worst subset perform compared to the full model? Model explosion when complete separation happens? Need to train up a central subset? Show confusion matrices and model parameter estimates.

```{r}
quick_classes <- pc_circled %>%
  select(feature, quick_class) %>%
  filter(quick_class!="Unclassified") %>%
  mutate(quick_class=quick_class=="Good")
summary(quick_classes)

full_quick_model <- FT2040_features %>% 
  filter(feature%in%quick_classes$feature) %>%
  select(-feat_class) %>%
  left_join(quick_classes, by="feature") %>%
  select(-feature) %>%
  glm(formula=quick_class~., family = binomial)

mini_quick_model <- FT2040_features %>% 
  filter(feature%in%quick_classes$feature) %>%
  select(-feat_class) %>%
  left_join(quick_classes, by="feature") %>%
  select(quick_class, med_cor, med_SNR) %>%
  glm(formula=quick_class~., family = binomial)

FT2040_features %>%
  mutate(pred_prob=predict(object=mini_quick_model, newdata=., type="response")) %>%
  mutate(pred_class=ifelse(pred_prob>0.5, "Good", "Bad")) %>%
  with(table(feat_class, pred_class))
```